---
title: "Hands-on Exercise 2"
author: "Jin Yao"
---

# Install and load required libraries

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse) 
```

## Importing the data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

###ggrepel - to repel overlapping text

replace `geom_text()` by [`geom_text_repel()`](https://ggrepel.slowkow.com/reference/geom_text_repel.html) and `geom_label()` by [`geom_label_repel`](https://ggrepel.slowkow.com/reference/geom_text_repel.html)

#### original plot

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

#### with ggrepel

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

## ggplot2 built in themes - example using theme_gray()

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_gray() +
  ggtitle("Distribution of Maths scores") 
```

### ggtheme package - example withe theme_economist()

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```

working with hrbthems package

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```
adjusting the size of bottom left Maths - axis_title size to set the size, base_size for the default axis labels and grid = "y" to remove x-axis grid lines 

```{r} 
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(axis_title_size = 18,
              base_size = 15,
              grid ="Y")
```

Additional online - another way to remove all the grid use - + theme(panel.background = element_blank())
```{r} 
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme(panel.background = element_blank())
```


To display more than 1 graph:scatterplot for English score versus Maths score

1st graph: math(addition of p1 before the code to facilitate patchwork extension later on)
```{r}
p1 <- ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") + 
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of Maths scores")
```


2nd graph English(addition of p2 before the code to facilitate patchwork extension later on)
```{r}
p2 <- ggplot(data=exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of English scores")
```

Combined graph (addition of p3 before the code to facilitate patchwork extension later on)
```{r}
p3 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y= ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

### Composite graphs using patchwork extension

in patchwork , + is a 2 column layout
```{r}
p1+p2
```
 Additional function of patchwork / and | and ()

```{r}
(p1 | p2)
```

```{r}
(p1 / p2)
```

combining three ggplot graphs 
```{r}
(p1 / p2) | p3
```
 

 
```{r}
p3/(p1|p2|p3)/(p1+p2)
```
 
 
inset element - put a small graph at the top or bottom of a bigger chart
```{r}
p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
``` 

composite figure with tagging for the different charts
```{r}
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')
```

tryout with different numbers 
```{r}
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'A')
```
 More plot annotation capability: title, subtitle, captions, tag levels plot tag control element size
```{r}
patchwork <- (p1 + p2) / p3
patchwork + plot_annotation(
  title = 'Trying out title for 3 graphs',
  subtitle = 'These 3 plots will reveal something about the data',
  caption = 'Disclaimer: I\'m just trying out',
  tag_levels = c('A','1')) & theme(plot.tag = element_text(size = 8))
``` 
```{r}
patchwork[[1]] <- patchwork[[1]] + plot_layout(tag_level = 'new')
patchwork + plot_annotation(tag_levels = c('A', '1'))
```


assigning operator pre-fix e.g adding fig at the front
```{r}
patchwork + plot_annotation(tag_levels = c('A', '1'), tag_prefix = 'Fig. ',
                            tag_sep = '.', tag_suffix = ':')
```





 using patchwork together with the theme
 
```{r} 
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
``` 

try out - theme part can insert other ggtheme learned before e.g theme_gray() etc

```{r} 
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
``` 
